---
title: "Early Glargine in DKA"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyr)
library(magrittr)
library(dplyr)
library(lubridate)
library(zoo) 
library(M3)
library(data.table)
library(Table1)
library(devtools)
library(Hmisc)
library(knitr)

# read in data
final <- read.csv("H:\\Endocrinology\\Ohman\\Ohman early glargine\\Data\\final data from sas.csv")

# use Claire's functions
source(file='C:\\Users\\pylell\\Documents\\GitHub\\General-code\\temp_table1_nofishers.r')


####################################################################################################

####################################################################################################
#Descriptive Tables #
####################################################################################################

# Set factors and levels
final$rebound_hyper <- as.factor(final$rebound_hyper)
final$group <- as.factor(final$group)
final$cerebral_edema <- as.factor(final$cerebral_edema)
levels(final$rebound_hyper)=c('No','Yes')
levels(final$Sex)=c('Female','Male')
levels(final$Ethnicity)=c('Black or African American','Hispanic','Multiracial','Other','White')
levels(final$Insurance_Type)=c('Unknown','Private/Military','Public','Self Pay')
final$hypo_12hr <- as.factor(final$hypo_12hr)
final$hypo_iv <- as.factor(final$hypo_iv)
final$ketosis_either <- as.factor(final$ketosis_either)
levels(final$hypo_12hr)=c('No','Yes')
levels(final$hypo_iv)=c('No','Yes')
levels(final$ketosis_either)=c('No','Yes')
levels(final$cerebral_edema)=c('No','Yes')
final$hypokal_12hr <- as.factor(final$hypokal_12hr)
levels(final$hypokal_12hr) =c('No','Yes')
final$hypokal_iv <- as.factor(final$hypokal_iv)
levels(final$hypokal_iv)=c('No','Yes')

# convert to numeric
final$HbA1c_Closest_to_Admisson <- as.numeric(final$HbA1c_Closest_to_Admisson)

# Set Labels
label(final$rebound_hyper)='Rebound hyperglycemia'
label(final$group)="Group"
label(final$Sex)='Sex'
label(final$Ethnicity)='Ethnicity'
label(final$Insurance_Type)='Insurance'
label(final$Age)='Age'
label(final$Height)='Height'
label(final$Weight)='Weight'
label(final$HbA1c_Closest_to_Admisson)='HbA1c closest to admission'
label(final$Median_Income)='Median income'
label(final$median_glucose)='Median glucose'
label(final$hypo_12hr)='Hypoglycemia within 12 hours after d/c IV insulin'
label(final$hypo_iv)='Hypoglycemia during IV insulin'
label(final$ketosis_either)='Ketosis by either BHB or urine ketones'
label(final$cerebral_edema)='Cerebral Edema'
label(final$hypokal_12hr)='Hypokalemia within 12 hours after d/c IV insulin'
label(final$hypokal_iv)='Hypokalemia during IV insulin'
label(final$dur_iv_insulin)='Length of IV insulin administration'
label(final$slopegluc)='Rate of glucose decrease (mg/dL per minute)'

Norig <- length(final$group)
final <- final[final$group %in% c("Early","Late"),]
final$group <- droplevels(final$group)
N <- sum(!is.na(final$group))
Nexcl <- Norig - N

tab1 <- final_table(final,c('Sex','Ethnicity',"Insurance_Type","Age","Height","Weight","HbA1c_Closest_to_Admisson","Median_Income"),
                    group=final$group,margin=2,single=0,ron=2)


####################################################################################################

####################################################################################################
#Primary Outcome #
####################################################################################################

# early vs. late glargine variable is October.1.2014.Flag
#tbl <- table(rebound_hyper$rebound_hyper,rebound_hyper$October.1.2014.Flag)
#rslt <- chisq.test(tbl)
tab_primary <- final_table(final,c('rebound_hyper','median_glucose','hypo_12hr','hypo_iv','ketosis_either','cerebral_edema',
                                   'hypokal_12hr','hypokal_iv','dur_iv_insulin','slopegluc'),
                           group=final$group,margin=2,single=0,ron=2)



```

# Background

The purpose of this study was to determine if early glargine administration in children with DKA prevents rebound hyperglycemia and recurrent ketogenesis without increasing the rate of hypoglycemia.   

# Methods

Early glargine administration was defined as either (1) glargine administered before the start of IV insulin administration or (2) glargine administered after the start of IV insulin administration with >4 hours of overlap with IV insulin.  Late glargine administration was defined as glargine administered with <2 hour overlap with IV insulin.  There were `r Nexcl` participants with between 2-4 hours of overlap with IV insulin who were excluded from this analysis.
 
The primary outcome of this study is rebound hyperglycemia, defined as serum glucose >180 mg/dL within 12 hours of discontinuation of IV insulin.  

The following data cleaning and manipulation steps were performed in order to determine whether each participant met the definition of rebound hyperglycemia:

1)  Study IDs on the main data sheet ending in "_1" or something similar, to indicate distinct encounters, were stripped of the suffix starting with the underscore, so they would match the IDs on the glucose data sheet.
2)  Missing glucose values or those with comments such as ("Already Documented","See Comment","Task Duplication", and "TNP") deleted from the glucose results.
3)  Encounters that were missing insulin administration dates and/or times were deleted.
4)  For each encounter, the last date and time of IV insulin administration was determined.
5)  The main data sheet was merged with glucose results by a combination of study ID and age, since a participant could have more than one encounter.
6)  If a participant had a glucose value >180 mg/dL within 12 hours of IV insulin discontinuation, s/he was classified as having rebound hyperglycemia.

Additional data cleaning and manipulation steps included:

1)  Military and private insurance were combined into one category.
2)  American Indian and Asian ethnicities were combined with Other ethnicity.

Secondary outcomes were defined as follows:

1) If a participant had a glucose value <70 within 12 hours of IV insulin discontinuation, s/he was classified as having hypoglycemia within 12 hours of insulin discontinuation.
2) If a participant had a glucose value <70 during IV insulin administration, s/he was classified as having hypoglycemia during IV insulin.
3) If a participant had a beta-hydroxybutyrate value >= 1.5 mmol/L or moderate or large urine ketones within 12 hours of IV insulin discontinuation, s/he was classifed as having ketosis within 12 hours of insulin discontinuation.
4) If a participant had a potassium level <3.5 mmol/L either during IV insulin administration or within 12 hours of IV insulin discontinuation, s/he was classified as having hypokalemia during that time period.

Descriptive statistics reported include mean and standard deviations for continuous variables, and frequencies and percentages for categorical variables.  To compare characteristics in the two groups, t-tests were used for continuous variables and the chi-square test was used for categorical variables.

# Results

There were `r N` subjects included in this analysis.  Descriptive statistics are shown in Table 1.

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab1,caption="Table 1.  Descriptive statistics.")
```
\

Table 2 shows the results of the comparisons of study outcomes.

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab_primary,caption="Table 2.  Comparison of study outcomes.")
```



